---
title: "Genomics of Drug Sensitivity in Cancer (GDSC) Project"
author: "Sam Scott, Megan Lee, Titus Sun"
date: "2026-06-01"
output: html_document
---

```{r warning=FALSE, message=FALSE}
library(caret)
library(tidyverse)
library(kernlab)
library(dplyr)
library(tibble)
library(mlbench)
library(randomForest)
library(corrplot)
library(ggplot2)
#install.packages("patchwork", repos = "https://cloud.r-project.org")
#url("https://cloud.r-project.org")
#options(download.file.method = "wininet")
#install.packages("patchwork", repos = "https://cloud.r-project.org")
library(patchwork)
library(Hmisc)
library(e1071)
library(doParallel)
library(kknn)
library(earth)
```

# a way to read the dataset off of github so no need for local machine calls
```{r}
data_url <- "https://raw.githubusercontent.com/sascott9655/GDSC-Project/refs/heads/main/GDSC_DATASET.csv"
data  <- read.csv(data_url)
```

```{r}
str(data)
head(data)
dim(data) #242,035 entries, 19 total columns
```

```{r}
# cleaning the data

# verifying the COSMIC_ID numbers are unique to each CELL_LINE_NAME

data %>%
  group_by(COSMIC_ID) %>%
  summarise(n_unique = n_distinct(CELL_LINE_NAME)) %>%
  filter(n_unique > 1)

# verifying the DRUG_ID numbers are unique to each DRUG_NAME

data %>%
  group_by(DRUG_ID) %>%
  summarise(n_unique = n_distinct(DRUG_NAME)) %>%
  filter(n_unique > 1)

# assessing the redundancy of the TCGA columns

tcga_match <- data$TCGA_DESC == data$Cancer.Type..matching.TCGA.label. #see if TCGA_DESC has values that are the same as Cancer Type
table(tcga_match, useNA = "ifany")

sum(data$Cancer.Type..matching.TCGA.label. == "") # see all Cancer Types that are empty strings (51446)
sum(data$TCGA_DESC == "") # see all TCGA_DESC that are empty strings (1067)
sum(data$Cancer.Type..matching.TCGA.label. == "" & # see all where both columns are empty (707)
      data$TCGA_DESC == "")

# removing irrelevant and redundant columns
data2 <- data %>%
  select(-COSMIC_ID, -DRUG_ID, -Screen.Medium, -Growth.Properties, -CNA, -Gene.Expression, -Methylation)
str(data2)
```

```{r}
# looking for na and blank values

data2 %>%
  summarise(across(everything(), ~sum(is.na(.))))%>%
  t()

#show what values have empty values in a tibble called blank

blank <- sapply(data2, function(x) sum(x == "", na.rm = TRUE))
as_tibble(list(
  column = names(blank),
  empty_count = as.vector(blank)
))
```

```{r}
# reference Cancer.Type..matching.TCGA.label. to enter values for TCGA_DESC

unique(data2$Cancer.Type..matching.TCGA.label.) #Cancer Types unique values

#Essentially if empty values occur in both Cancer Type and TCGA_DESC, label it as
#UNCLASSIFIED1. Otherwise use the TCGA_DESC value.

data2 <- data2 %>% 
  mutate(
    TCGA_DESC = case_when(
      TCGA_DESC == "" & Cancer.Type..matching.TCGA.label. == "" ~ "UNCLASSIFIED1",
      TCGA_DESC == "" ~ Cancer.Type..matching.TCGA.label.,
      TRUE ~ TCGA_DESC
    )
  )

#Now there are no empty string values for our TCGA_DESC

blank <- sapply(data2, function(x) sum(x == "", na.rm = TRUE))
as_tibble(list(
  column = names(blank),
  empty_count = as.vector(blank)
))
unique(data2$TCGA_DESC)
```

```{r}
# reference TCGA_DESC to enter GDSC.Tissue.descriptor.1 values for blank values in GDSC.Tissue.descriptor.2

length(unique(data2$GDSC.Tissue.descriptor.1))
length(unique(data2$GDSC.Tissue.descriptor.2))
length(unique(data2$TCGA_DESC))


#Find non-empty or non-missing GDSC descriptor 1 values
#Count the number of pairs of occurrences
#Keep most common descriptor values
#Return those common mapped descriptor values

gdsc_miss <- data2 %>%
  filter(GDSC.Tissue.descriptor.1 != "", !is.na(GDSC.Tissue.descriptor.1)) %>%
  group_by(TCGA_DESC, GDSC.Tissue.descriptor.1) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(TCGA_DESC) %>%
  slice_max(n, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(TCGA_DESC, GDSC.Tissue.descriptor.1)

#rename gdsc.Tissue.descriptor to gdsc_map

gdsc_miss <- gdsc_miss %>%
  rename(gdsc_map = GDSC.Tissue.descriptor.1)

#Add gdsc_miss to our dataframe data2
#Use GDSC Tissue descriptor 2 if not empty
#Otherwise use GDSC Tissue descriptor 1 
#Otherwise use gdsc_map (the most common descriptor per TCGA_DESC)
#Otherwise "UNCLASSIFIED1"

data2 <- data2 %>%
  left_join(gdsc_miss, by = "TCGA_DESC") %>%
  mutate(
    GDSC.Tissue.descriptor.2 = case_when(
      GDSC.Tissue.descriptor.2 != "" ~ GDSC.Tissue.descriptor.2,
      GDSC.Tissue.descriptor.1 != "" ~ GDSC.Tissue.descriptor.1,
      !is.na(gdsc_map) ~ gdsc_map,
      TRUE ~ "UNCLASSIFIED1"
    )
  )

#No empty or missing values for GDSC Tissue Descriptor 2

blank <- sapply(data2, function(x) sum(x == "", na.rm = TRUE))
as_tibble(list(
  column = names(blank),
  empty_count = as.vector(blank)
))
unique(data2$GDSC.Tissue.descriptor.2)
```

```{r}
# imputing blank MSI values

unique(data2$Microsatellite.instability.Status..MSI.)

tcga_msi <- data2 %>%
  group_by(TCGA_DESC) %>%
  summarise(
    n_groups = n_distinct(Microsatellite.instability.Status..MSI.),
    groups = paste(unique(Microsatellite.instability.Status..MSI.), collapse = ", "),
    .groups = "drop"
  )
tcga_msi

table(data2$Microsatellite.instability.Status..MSI.)

#Find non-empty or non-missing msi values
#Count the number of pairs of occurrences
#Keep most common MSI values
#Return those common mapped MSI values

msi_miss <- data2 %>%
  filter(Microsatellite.instability.Status..MSI. != "", !is.na(Microsatellite.instability.Status..MSI.)) %>%
  group_by(TCGA_DESC, Microsatellite.instability.Status..MSI.) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(TCGA_DESC) %>%
  slice_max(n, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(TCGA_DESC, Microsatellite.instability.Status..MSI.)

#rename Microsatellite.instability.Status..MSI. to msi_map

msi_miss <- msi_miss %>%
  rename(msi_map = Microsatellite.instability.Status..MSI.)

#Add msi_miss to our dataframe data2
#Make empty string NA values for the coalesce function to work
#Prioritize using the MSI values if present 
#Secondly use msi_map (the most common descriptor per TCGA_DESC, the imputation part of the process)
#Otherwise "UNCLASSIFIED1"

data2 <- data2 %>%
  left_join(msi_miss, by = "TCGA_DESC") %>%
  mutate(
    Microsatellite.instability.Status..MSI. = coalesce(
      na_if(Microsatellite.instability.Status..MSI., ""),
      msi_map,
      "UNCLASSIFIED1"
    )
  )

#No more missing or empty MSI values

blank <- sapply(data2, function(x) sum(x == "", na.rm = TRUE))
as_tibble(list(
  column = names(blank),
  empty_count = as.vector(blank)
))
unique(data2$Microsatellite.instability.Status..MSI.)
```

```{r}
# imputing blank TARGET values

sum(data2$TARGET == "")
sum(data2$TARGET == "" & data2$TARGET_PATHWAY != "Unclassified")
sum(data2$TARGET_PATHWAY == "")

target_miss <- data2 %>%
  filter(TARGET != "", TARGET != " ", !is.na(TARGET)) %>%
  group_by(TARGET_PATHWAY, TARGET) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(TARGET_PATHWAY) %>%
  slice_max(n, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(TARGET_PATHWAY, TARGET)

target_miss <- target_miss %>%
  rename(target_map = TARGET)

names(target_miss)
names(data2 %>% left_join(target_miss, by = "TARGET_PATHWAY"))

data2 <- data2 %>%
  left_join(target_miss, by = "TARGET_PATHWAY") %>%
  mutate(
    TARGET = coalesce(
      na_if(str_trim(TARGET), ""),
      target_map,
      "UNCLASSIFIED1"
    )
  )

sum(trimws(data2$TARGET) == "", na.rm = TRUE)
blank <- sapply(data2, function(x) sum(x == "", na.rm = TRUE))
as_tibble(list(
  column = names(blank),
  empty_count = as.vector(blank)
))
unique(data2$TARGET)
```

```{r}
# data_c is the cleaned version of the GDSC data frame.

# making sure our data_c has no empty or missing values in predictors

str(data2)
data_c <- data2 %>%
  select(-GDSC.Tissue.descriptor.1, -Cancer.Type..matching.TCGA.label., -gdsc_map, -msi_map, -target_map)
str(data_c)

unique(data_c$TARGET)

blank_c <- sapply(data_c, function(x) sum(x == "", na.rm = TRUE))
as_tibble(list(
  column = names(blank_c),
  empty_count = as.vector(blank_c)
)
)

sum(data_c$TARGET == " ")
```

```{r}
# correlation matrix with numeric data
data_num <- data_c[, sapply(data_c, is.numeric)] #3 out of the 10 predictors are numeric 
data_corr <- cor(data_num)
?cor
corr_mat <- cor(data_num, use = "complete.obs") #make sure that no NAs values are present when doing correlations
corr_mat
corrplot(data_corr)
```

```{r}
# boxplots of MSI status

msi_ln <- ggplot(data_c, aes(x = Microsatellite.instability.Status..MSI., y = LN_IC50)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "MSI Status", y = "LN_IC50",
       title = "LN_IC50 across MSI groups")

msi_auc <- ggplot(data_c, aes(x = Microsatellite.instability.Status..MSI., y = AUC)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "MSI Status", y = "AUC",
       title = "AUC across MSI groups")

msi_z <- ggplot(data_c, aes(x = Microsatellite.instability.Status..MSI., y = Z_SCORE)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "MSI Status", y = "Z_SCORE",
       title = "Z_SCORE across MSI groups")

(msi_ln + msi_auc + msi_z) & theme(aspect.ratio = 1)
```

```{r}
# boxplots of TARGET

unique(data_c$TARGET_PATHWAY)

tp_ln <- ggplot(data_c, aes(x = TARGET_PATHWAY, y = LN_IC50)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "Target Pathway", y = "LN_IC50",
       title = "LN_IC50 across Target Pathways")

tp_auc <- ggplot(data_c, aes(x = TARGET_PATHWAY, y = AUC)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "Target Pathway", y = "AUC",
       title = "AUC across Target Pathways")

tp_z <- ggplot(data_c, aes(x = TARGET_PATHWAY, y = Z_SCORE)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "Target Pathway", y = "Z_SCORE",
       title = "Z_SCORE across Target Pathways")

(tp_ln / tp_auc / tp_z) +
  plot_layout(heights = c(3, 3, 3)) &
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

```{r}
# boxplots of TCGA_DESC

unique(data_c$TCGA_DESC)

tcga_ln <- ggplot(data_c, aes(x = TCGA_DESC, y = LN_IC50)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "Cancer Type", y = "LN_IC50",
       title = "LN_IC50 across Cancer Types")

tcga_auc <- ggplot(data_c, aes(x = TCGA_DESC, y = AUC)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "Cancer Type", y = "AUC",
       title = "AUC across Cancer Types")

tcga_z <- ggplot(data_c, aes(x = TCGA_DESC, y = Z_SCORE)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "Cancer Type", y = "Z_SCORE",
       title = "Z_SCORE across Cancer Types")

(tcga_ln / tcga_auc / tcga_z) +
  plot_layout(heights = c(3, 3, 3)) &
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

```{r}
# histogram of numerical variables

par(mfrow = c(3, 1))
hist.data.frame(data_num)
```

```{r}
# looking at the histograms, AUC requires a transformation.

skewness(data_c$AUC)

auc_log <- data_c$AUC
auc_log <- pmin(pmax(auc_log, 0.001), 0.999) #log values 0 and 1 are undefined so we need to set a minimum and maximum
auc_trans <- log(auc_log / (1 - auc_log))
skewness(auc_trans)
```

```{r}
# centering and scaling new data frame of numerical predictor data with transformed values

data_numpred_trans <- data.frame(
  AUC_trans = auc_trans,
  Z_SCORE = data_c$Z_SCORE
)

par(mfrow = c(2, 1)) #draws 2 plots side by side 2 rows by 1 column
hist.data.frame(data_numpred_trans)

data_numpred_cs <- scale(data_numpred_trans)
```

```{r}
# principal components analysis

data_num_pca <- prcomp(data_numpred_cs)
data_num_pca

#how variance each PC contributes
exp_var <- (data_num_pca$sdev^2) / sum(data_num_pca$sdev^2)
exp_var

#cumulatively adds how much each PC contributes till the value 1 is reached
data_ex_vratio <- cumsum(exp_var)
data_ex_vratio 

par(pty = "s") #generate a square plot region s for square
plot(
  exp_var,
  type = "b", #draw both lines and points
  pch = 19, #use solid circles for the points
  xlab = "Principal Component",
  ylab = "Explained Variance Ratio",
  main = "Scree Plot"
)

abs(data_num_pca$rotation[,1])
sort(abs(data_num_pca$rotation[,1]), decreasing = TRUE)

biplot(data_num_pca)

PC1 <- data_num_pca$x[,1]
cor.test(PC1, data_c$LN_IC50) #show PC 1 confidence interval roughly between 0.746 and 0.750

plot(
  PC1,
  data_c$LN_IC50,
  pch = 19,
  xlab = "PC1",
  ylab = "LN_IC50",
  main = "PC1 vs LN_IC50"
)
abline(lm(data_c$LN_IC50 ~ PC1), col = "red")
```

```{r}
# the final cleaned, reduced, and transformed data set is data_f.
data_f <- data_c %>% select(-AUC)
str(data_f)
```

#SAM

#Since the models took an extremely long time to train due to the caret package
#naturally encoding all your categorical variables to be dummy variables, it was computationally expensive.
#Instead predictors like CELL_LINE_NAME and DRUG_NAME are removed from the model to decrease computation time
#and they did not provide predictive power


```{r} 
data_f <- data_f %>% select(-CELL_LINE_NAME, -DRUG_NAME)
```

#Seed is set. Data observations amount are reduced for faster training.

```{r}
seed <- 756
set.seed(seed)

#Target variable
IC50 <- data_f$LN_IC50

#Predictors
data_f <- as.data.frame(data_f)

row_reduction <- 0.10

#Get 10 percent of the data amount randomly
rrow_count <- sample(nrow(data_f), size=floor(row_reduction*nrow(data_f)))

#Reduced the the amount of observations for modeling
data_f <- data_f[rrow_count, ]
IC50 <- IC50[rrow_count]
```

#Encode -- One hot encode all the predictor values manually (not through caret model)
#that will be used for our model predictions

```{r}
#Encoding TARGET variables
target_encode <- model.matrix(~ TARGET - 1, data = data_f)

#Encoding the TARGET_PATHWAY variables
pathway_encode <- model.matrix(~ TARGET_PATHWAY - 1, data = data_f)

#Encoding TCGA_DESC
tissue_encode <- dummyVars(~ TCGA_DESC, data = data_f)
tissue_encode <- predict(tissue_encode, data_f)

#Encoding GDSC Tissue Descriptor 2
gdsctd2_encode <- dummyVars(~ GDSC.Tissue.descriptor.2, data = data_f)
gdsctd2_encode <- predict(gdsctd2_encode, data_f)

#Since only 3 types in predictor we can make it discrete numeric 
data_f$MSI <- as.numeric(as.factor(data_f$Microsatellite.instability.Status..MSI.))
```

#Combine all encoded variables 

```{r}
predictors <- cbind(
  target_encode,
  pathway_encode,
  tissue_encode,
  gdsctd2_encode,
  MSI = data_f$MSI,
  Z_SCORE = data_f$Z_SCORE
)
predictors <- as.data.frame(predictors)
```


#Train test split using stratified sampling

```{r}
trainingRows <- createDataPartition(IC50, p =.8, list=FALSE)

trainPredictors <- predictors[trainingRows,]
testPredictors <- predictors[-trainingRows,]

trainIC50 <- IC50[trainingRows]
testIC50 <- IC50[-trainingRows]
```

#Encode variables that are highly correlated and provide zero variance are removed

```{r}
#Get rid of variables that are zero variance...
nzv <- nearZeroVar(trainPredictors)
trainPredictors <- trainPredictors[, -nzv]
testPredictors  <- testPredictors[, -nzv]

#...and highly correlated
correlations <- cor(trainPredictors)
highCorr <- findCorrelation(correlations, cutoff = 0.75)
trainPredictors <- trainPredictors[, -highCorr]
testPredictors  <- testPredictors[, -highCorr] # Remove same columns 
```

```{r}
colnames(trainPredictors)
```

```{r}
#Before model training we need to use parallel processing so the CPU can run models more quickly
num_cores <- detectCores() - 1             # Leave 1 core for background tasks
cl <- makeCluster(num_cores)
registerDoParallel(cl)

set.seed(seed)

#Set up cross validation 
fivefoldCV <- trainControl(method = "cv", number = 5, allowParallel = TRUE)
#repeatedSplits <- trainControl(method = "repeatedcv", number = 5, repeats = 3, allowParallel = TRUE)
```

```{r}
#Sanity Check
dim(trainPredictors)
dim(testPredictors)
length(trainIC50)
length(testIC50)
```

#Linear Regression

```{r}
set.seed(seed)
lin_reg1 <- train(trainPredictors, trainIC50,
                 method = "lm",
                 trControl = fivefoldCV)
lin_reg1
```

```{r}
#set.seed(seed)
#lin_reg2 <- train(trainPredictors, trainIC50,
#                 method = "lm",
#                 trControl = repeatedSplits)
#lin_reg2
```

#KNN regression

```{r}
set.seed(seed)
knn <- train(trainPredictors, trainIC50,
             method = "kknn",
             tuneGrid = data.frame(
             kmax = seq(1, 21, by = 2), 
             distance = c(2),            
             kernel = c("rectangular")           
             ),
             preProcess = c("center", "scale", "pca"),
             trControl = fivefoldCV)
knn
```

```{r}
#set.seed(seed)
#knn2 <- train(trainPredictors, trainIC50,
#             method = "kknn",
#             tuneGrid = data.frame(
#             kmax = seq(2, 12, by = 2), 
#             distance = 2,            
#             kernel = "rectangular"           
#             ),
#             preProcess = c("center", "scale", "pca"),
#             trControl = repeatedSplits)
#knn2
```

#SVM

```{r}
set.seed(seed)
SVM <- train(trainPredictors, trainIC50,
             method = "svmRadial",
             tuneLength = 5,
             preProcess = c("center", "scale", "pca"),
             trControl = fivefoldCV) 
SVM
```

```{r}
#set.seed(seed)
#SVM2 <- train(trainPredictors, trainIC50,
#             method = "svmLinear",
#             preProcess = c("center", "scale", "pca"),
#             trControl = repeatedSplits) 
#SVM2
```


#MARS

```{r}
set.seed(seed)
MARSGrid <- expand.grid(degree = 1:2, nprune = 2:30)
MARS <- train(trainPredictors, trainIC50,
              method ="earth",
              tuneGrid = MARSGrid,
              trControl = fivefoldCV)
MARS
```

```{r}
#set.seed(seed)
#MARSGrid2 <- expand.grid(degree = 1:4, nprune = 2:20)
#MARS2 <- train(trainPredictors, trainIC50,
#              method ="earth",
#              tuneGrid = MARSGrid2,
#              trControl = repeatedSplits)
#MARS2
```

#Random Forest

```{r}
max_mtry <- ncol(trainPredictors)
max_mtry
```

```{r}
mtryGrid <- data.frame(mtry = floor(seq(2, ncol(trainPredictors), length = 10)))
set.seed(seed)
rf <- train(trainPredictors, trainIC50,
                method = "rf",
                tuneGrid = mtryGrid,
                ntree = 750,
                importance = TRUE,
                trControl = fivefoldCV)
rf
```

```{r}
#mtryGrid2 <- data.frame(mtry = floor(seq(10, ncol(trainPredictors)/2, length = 15)))

#set.seed(seed)
#rf2 <- train(trainPredictors, trainIC50,
#                method = "rf",
#                tuneGrid = mtryGrid2,
#                ntree = 500,
#                importance = TRUE,
#                trControl = repeatedSplits)
#rf2
```


```{r}
#test our model predictions
testResults_lin_regr <- predict(lin_reg1, testPredictors)
testResults_knn_regr <- predict(knn, testPredictors)
testResults_SVM <- predict(SVM, testPredictors)
testResults_MARS <- predict(MARS, testPredictors)
testResults_rf <- predict

#Use postSample to evaluate the RMSE, Rsquared, and MAE
lin_regr_metrics  <- postResample(testResults_lin_regr, testIC50)
knn_regr_metrics  <- postResample(predict(knn, testPredictors), testIC50)
SVM_metrics  <- postResample(predict(SVM, testPredictors), testIC50)
MARS_metrics <- postResample(predict(MARS, testPredictors), testIC50)
rf_metrics   <- postResample(predict(rf, testPredictors), testIC50)

#model results on y_test 
model_results <- cbind(
  Linear_Regression = lin_regr_metrics,
  Knn_Regression = knn_regr_metrics,
  SVM = SVM_metrics,
  MARS = MARS_metrics,
  Random_Forest = rf_metrics
)
model_results
```







#It would cool to graph how affecting tuning parameters can change the outcome of
#models. I did it for a different SVM model down here but this is just some ideas.



#Choosing Z-Score as a predictor to look at to see how the tuning parameters impact SVM model

```{r}
z_score_line <- tibble(Z_SCORE = seq(min(trainPredictors$Z_SCORE),
                                     max(trainPredictors$Z_SCORE),
                                     length.out = 200))
#z_score_line
```

```{r}
SVM_explore <- function(cost, epsilon, 
                        data_grid = z_score_line, train_data = trainPredictors,
                        train_y = trainIC50) {
  
svmFit <- ksvm(
  x = as.matrix(train_data$Z_SCORE),
  y = train_y,
  kernel = "rbfdot",
  C = cost,
  epsilon = epsilon)

predictions <- predict(svmFit, as.matrix(data_grid$Z_SCORE))

tibble(
  Z_SCORE = data_grid$Z_SCORE,
  IC50_pred = predictions,
  cost = cost,
  epsilon = epsilon)

}

```

#Changes that occur to SVM model when cost changes

```{r}
cost_varies <- 10^seq(-2,1, length = 4) #change in cost (0.01, 0.10, 1.00, 10.00)
eps01_fixed <- rep(0.1, 4) #4 iterations of the same epsilon value 0.1

cost_curves <- map2_dfr(cost_varies,
                        eps01_fixed,
                        ~ SVM_explore(.x, .y))
ggplot(cost_curves, aes(x = Z_SCORE, y = IC50_pred, color = factor(cost))) + geom_line(size=1)
```

#Changes that occur to SVM model when epsilon changes

```{r}
eps_varies <- c(0.01, 0.1, 0.5, 1.0) #epsilon values vary
cost_fixed <- rep(0.1, 4) #cost values are fixed
eps_curves <- map2_dfr(cost_fixed,
                        eps_varies,
                        ~ SVM_explore(.x, .y))
ggplot(eps_curves, aes(x = Z_SCORE, y = IC50_pred, color = factor(epsilon))) + geom_line(size=1)
```



```{r}
#Stop the parallel cluster when models are finished training
stopCluster(cl)
registerDoSEQ()
```