---
title: "Appendix 1. Genomics of Drug Sensitivity in Cancer (GDSC) Project"
author: "Sam Scott, Megan Lee, Titus Sun"
date: "2026-06-01"
output: github_document
---

```{r warning=FALSE, message=FALSE}
library(caret)
library(tidyverse)
library(kernlab)
library(dplyr)
library(tibble)
library(mlbench)
library(randomForest)
library(corrplot)
library(ggplot2)
#install.packages("patchwork", repos = "https://cloud.r-project.org")
#url("https://cloud.r-project.org")
#options(download.file.method = "wininet")
#install.packages("patchwork", repos = "https://cloud.r-project.org")
library(patchwork)
library(Hmisc)
library(e1071)
library(doParallel)
library(kknn)
library(earth)
library(rcompanion)
library(gridExtra)
library(grid)
```

# a way to read the dataset off of github so no need for local machine calls
```{r}
data_url <- "https://raw.githubusercontent.com/sascott9655/GDSC-Project/refs/heads/main/GDSC_DATASET.csv"
data  <- read.csv(data_url)
```

```{r}
str(data)
#head(data)
dim(data) # 242,035 entries, 19 total columns
colnames(data)

# verifying that each cell line was tested once with each drug

data %>%
  count(CELL_LINE_NAME, sort = TRUE)
data %>%
  count(DRUG_NAME, sort = TRUE)
```

```{r}
# cleaning the data

# verifying the COSMIC_ID numbers are unique to each CELL_LINE_NAME

data %>%
  group_by(COSMIC_ID) %>%
  summarise(n_unique = n_distinct(CELL_LINE_NAME)) %>%
  filter(n_unique > 1)

# verifying the DRUG_ID numbers are unique to each DRUG_NAME

data %>%
  group_by(DRUG_ID) %>%
  summarise(n_unique = n_distinct(DRUG_NAME)) %>%
  filter(n_unique > 1)

# assessing the redundancy of the TCGA columns
# (To see if TCGA_DESC has values that are the same as Cancer Type.)

tcga_match <- data$TCGA_DESC == data$Cancer.Type..matching.TCGA.label. 
table(tcga_match, useNA = "ifany")

sum(data$Cancer.Type..matching.TCGA.label. == "") # see all Cancer Types that 
                                                  # are empty strings (51446)
sum(data$TCGA_DESC == "") # see all TCGA_DESC that are empty strings (1067)
sum(data$Cancer.Type..matching.TCGA.label. == "" & 
      data$TCGA_DESC == "") # see all where both columns are empty (707)

# removing irrelevant and redundant columns
data2 <- data %>%
  select(-COSMIC_ID, -DRUG_ID, -Screen.Medium, -Growth.Properties, -CNA, -Gene.Expression, -Methylation)
str(data2)
```

```{r}
# looking for na and blank values

data2 %>%
  summarise(across(everything(), ~sum(is.na(.))))%>%
  t()

#show what values have empty values in a tibble called blank

blank <- sapply(data2, function(x) {sum(x == "" | x == " ", na.rm = TRUE)})
as_tibble(list(
  column = names(blank),
  empty_count = as.vector(blank)
))
```

```{r}
# reference Cancer.Type..matching.TCGA.label. to enter values for TCGA_DESC

n_distinct(data2$TCGA_DESC)
n_distinct(data2$Cancer.Type..matching.TCGA.label.)
unique(data2$TCGA_DESC)
unique(data2$Cancer.Type..matching.TCGA.label.) # Cancer Types unique values

# Essentially if empty values occur in both Cancer.Type and TCGA_DESC, label it 
# as UNCLASSIFIED1. Otherwise use the TCGA_DESC value.

data2 <- data2 %>% 
  mutate(
    TCGA_DESC = case_when(
      TCGA_DESC == "" & Cancer.Type..matching.TCGA.label. == "" ~ "UNCLASSIFIED1",
      TCGA_DESC == "" ~ Cancer.Type..matching.TCGA.label.,
      TRUE ~ TCGA_DESC
    )
  )

#Now there are no empty string values for our TCGA_DESC

blank <- sapply(data2, function(x) {sum(x == "" | x == " ", na.rm = TRUE)})
as_tibble(list(
  column = names(blank),
  empty_count = as.vector(blank)
))
unique(data2$TCGA_DESC)
```

```{r}
# reference TCGA_DESC to enter GDSC.Tissue.descriptor.1 values for blank values 
# in GDSC.Tissue.descriptor.2

length(unique(data2$GDSC.Tissue.descriptor.1))
length(unique(data2$GDSC.Tissue.descriptor.2))
length(unique(data2$TCGA_DESC))

# Find non-empty or non-missing GDSC descriptor 1 values
# Count the number of pairs of occurrences
# Keep most common descriptor values
# Return those common mapped descriptor values

gdsc_miss <- data2 %>%
  filter(GDSC.Tissue.descriptor.1 != "", !is.na(GDSC.Tissue.descriptor.1)) %>%
  group_by(TCGA_DESC, GDSC.Tissue.descriptor.1) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(TCGA_DESC) %>%
  slice_max(n, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(TCGA_DESC, GDSC.Tissue.descriptor.1)

# rename gdsc.Tissue.descriptor1 to gdsc_map

gdsc_miss <- gdsc_miss %>%
  rename(gdsc_map = GDSC.Tissue.descriptor.1)

# Add gdsc_miss to our dataframe data2
# Use GDSC Tissue descriptor 2 if not empty
# Otherwise use GDSC Tissue descriptor 1 
# Otherwise use gdsc_map (the most common descriptor per TCGA_DESC)
# Otherwise "UNCLASSIFIED1"

data2 <- data2 %>%
  left_join(gdsc_miss, by = "TCGA_DESC") %>%
  mutate(
    GDSC.Tissue.descriptor.2 = case_when(
      GDSC.Tissue.descriptor.2 != "" ~ GDSC.Tissue.descriptor.2,
      GDSC.Tissue.descriptor.1 != "" ~ GDSC.Tissue.descriptor.1,
      !is.na(gdsc_map) ~ gdsc_map,
      TRUE ~ "UNCLASSIFIED1"
    )
  )

#No empty or missing values for GDSC Tissue Descriptor 2

blank <- sapply(data2, function(x) {sum(x == "" | x == " ", na.rm = TRUE)})
as_tibble(list(
  column = names(blank),
  empty_count = as.vector(blank)
))
unique(data2$GDSC.Tissue.descriptor.2)
```

```{r}
# imputing blank MSI values

unique(data2$Microsatellite.instability.Status..MSI.)

tcga_msi <- data2 %>%
  group_by(TCGA_DESC) %>%
  summarise(
    n_groups = n_distinct(Microsatellite.instability.Status..MSI.),
    groups = paste(unique(Microsatellite.instability.Status..MSI.), collapse = ", "),
    .groups = "drop"
  )
tcga_msi

table(data2$Microsatellite.instability.Status..MSI.)

# use chi-square test to check for stronger association of MSI with either
# TCGA_DESC or TARGET_PATHWAY for later grouping.

chisq.test(table(data2$Microsatellite.instability.Status..MSI., data2$TCGA_DESC))
chisq.test(table(data2$Microsatellite.instability.Status..MSI., data2$TARGET_PATHWAY))

# use Cramer's V to identify strength of associations

cramerV(table(data2$Microsatellite.instability.Status..MSI., data2$TCGA_DESC))
cramerV(table(data2$Microsatellite.instability.Status..MSI., data2$TARGET_PATHWAY))

# Find non-empty or non-missing msi values
# Count the number of pairs of occurrences
# Keep most common MSI values
# Return those common mapped MSI values

msi_miss <- data2 %>%
  filter(Microsatellite.instability.Status..MSI. != "", 
         !is.na(Microsatellite.instability.Status..MSI.)) %>%
  group_by(TCGA_DESC, Microsatellite.instability.Status..MSI.) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(TCGA_DESC) %>%
  slice_max(n, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(TCGA_DESC, Microsatellite.instability.Status..MSI.)

#rename Microsatellite.instability.Status..MSI. to msi_map

msi_miss <- msi_miss %>%
  rename(msi_map = Microsatellite.instability.Status..MSI.)

# Add msi_miss to our dataframe data2
# Make empty string NA values for the coalesce function to work
# Prioritize using the MSI values if present 
# Secondly use msi_map (the most common descriptor per TCGA_DESC, the 
# imputation part of the process)
# Otherwise "UNCLASSIFIED1"

data2 <- data2 %>%
  left_join(msi_miss, by = "TCGA_DESC") %>%
  mutate(
    Microsatellite.instability.Status..MSI. = coalesce(
      na_if(Microsatellite.instability.Status..MSI., ""),
      msi_map,
      "UNCLASSIFIED1"
    )
  )

#No more missing or empty MSI values

blank <- sapply(data2, function(x) {sum(x == "" | x == " ", na.rm = TRUE)})
as_tibble(list(
  column = names(blank),
  empty_count = as.vector(blank)
))
unique(data2$Microsatellite.instability.Status..MSI.)
```

```{r}
# imputing blank TARGET values

sum(data2$TARGET == "")
sum(data2$TARGET == "" & data2$TARGET_PATHWAY != "Unclassified")
sum(data2$TARGET_PATHWAY == "")

length(unique(data2$TARGET))
length(unique(data2$TARGET_PATHWAY))

target_miss <- data2 %>%
  filter(TARGET != "", TARGET != " ", !is.na(TARGET)) %>%
  group_by(TARGET_PATHWAY, TARGET) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(TARGET_PATHWAY) %>%
  slice_max(n, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(TARGET_PATHWAY, TARGET)

target_miss <- target_miss %>%
  rename(target_map = TARGET)

names(target_miss)
names(data2 %>% left_join(target_miss, by = "TARGET_PATHWAY"))

data2 <- data2 %>%
  left_join(target_miss, by = "TARGET_PATHWAY") %>%
  mutate(
    TARGET = coalesce(
      na_if(str_trim(TARGET), ""),
      target_map,
      "UNCLASSIFIED1"
    )
  )

#unique(data2$TARGET)

blank <- sapply(data2, function(x) {sum(x == "" | x == " ", na.rm = TRUE)})
as_tibble(list(
  column = names(blank),
  empty_count = as.vector(blank)
))
# unique(data2$TARGET)
```

```{r}
# data_c is the cleaned version of the GDSC data frame.

# making sure our data_c has no empty or missing values in predictors

str(data2)
data_c <- data2 %>%
  select(-GDSC.Tissue.descriptor.1, -Cancer.Type..matching.TCGA.label., -gdsc_map, -msi_map, -target_map)
str(data_c)

blank_c <- sapply(data_c, function(x) {sum(x == "" | x == " ", na.rm = TRUE)})
as_tibble(list(
  column = names(blank_c),
  empty_count = as.vector(blank_c)
)
)
```

```{r}
# boxplots of MSI status

msi_ln <- ggplot(data_c, aes(x = Microsatellite.instability.Status..MSI., y = LN_IC50)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "MSI Status", y = "LN_IC50",
       title = "LN_IC50 across MSI groups")

msi_auc <- ggplot(data_c, aes(x = Microsatellite.instability.Status..MSI., y = AUC)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "MSI Status", y = "AUC",
       title = "AUC across MSI groups")

msi_z <- ggplot(data_c, aes(x = Microsatellite.instability.Status..MSI., y = Z_SCORE)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "MSI Status", y = "Z_SCORE",
       title = "Z_SCORE across MSI groups")

(msi_ln + msi_auc + msi_z) &
  theme(
    aspect.ratio = 1,
    plot.margin = ggplot2::margin(10, 40, 10, 10),
    plot.title = element_text(size = 11, hjust = 0.5)
  )
```

```{r}
# boxplots of TARGET_PATHWAY

unique(data_c$TARGET_PATHWAY)

tp_ln <- ggplot(data_c, aes(x = TARGET_PATHWAY, y = LN_IC50)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "Target Pathway", y = "LN_IC50",
       title = "LN_IC50 across Target Pathways")

tp_auc <- ggplot(data_c, aes(x = TARGET_PATHWAY, y = AUC)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "Target Pathway", y = "AUC",
       title = "AUC across Target Pathways")

tp_z <- ggplot(data_c, aes(x = TARGET_PATHWAY, y = Z_SCORE)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "Target Pathway", y = "Z_SCORE",
       title = "Z_SCORE across Target Pathways")

(tp_ln / tp_auc / tp_z) +
  plot_layout(heights = c(3, 3, 3)) &
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

```{r}
# boxplots of TCGA_DESC

unique(data_c$TCGA_DESC)

tcga_ln <- ggplot(data_c, aes(x = TCGA_DESC, y = LN_IC50)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "Cancer Type", y = "LN_IC50",
       title = "LN_IC50 across Cancer Types")

tcga_auc <- ggplot(data_c, aes(x = TCGA_DESC, y = AUC)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "Cancer Type", y = "AUC",
       title = "AUC across Cancer Types")

tcga_z <- ggplot(data_c, aes(x = TCGA_DESC, y = Z_SCORE)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "Cancer Type", y = "Z_SCORE",
       title = "Z_SCORE across Cancer Types")

(tcga_ln / tcga_auc / tcga_z) +
  plot_layout(heights = c(3, 3, 3)) &
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

```{r}
# correlation matrix of numeric data

data_num <- data_c[, sapply(data_c, is.numeric)] # 3 out of the 10 predictors 
                                                 # are numeric 

data_corr <- cor(data_num)
corr_mat <- cor(data_num, use = "complete.obs") # make sure that no NAs values 
                                          # are present when doing correlations

corr_mat_tab <- tableGrob(head(round(corr_mat, 2)))
grid.newpage()
grid.draw(corr_mat_tab)
```

```{r}
# histogram of numerical variables

par(mfrow = c(3, 1))
hist.data.frame(data_num)
```

```{r}
# looking at the histograms, AUC requires a transformation.

skewness(data_c$LN_IC50)
skewness(data_c$AUC)
skewness(data_c$Z_SCORE)

auc_log <- data_c$AUC
auc_logc <- pmin(pmax(auc_log, 0.001), 0.999) # logit transformation requires
                                              # values strictly between 0 and 1.
auc_trans <- log(auc_logc / (1 - auc_logc))
skewness(auc_trans)
```

```{r}
# centering and scaling new data frame of numerical predictor data with 
# transformed values

data_numpred_trans <- data.frame(
  AUC_trans = auc_trans,
  Z_SCORE = data_c$Z_SCORE
)

par(mfrow = c(2, 1))
hist.data.frame(data_numpred_trans)

data_numpred_cs <- scale(data_numpred_trans) # centers and scales
```

```{r}
# principal components analysis

data_num_pca <- prcomp(data_numpred_cs)
data_num_pca

pca_table <- round(data_num_pca$rotation, 4)
grid.newpage()
grid.draw(tableGrob(pca_table))

abs(data_num_pca$rotation[,1])
sort(abs(data_num_pca$rotation[,1]), decreasing = TRUE)

biplot(data_num_pca)

# how variance each PC contributes
exp_var <- (data_num_pca$sdev^2) / sum(data_num_pca$sdev^2)
exp_var

# cumulatively adds how much each PC contributes till the value 1 is reached
data_ex_vratio <- cumsum(exp_var)
data_ex_vratio 

# Summary table
pca_summary <- data.frame(
  PC = paste0("PC", 1:length(data_num_pca$sdev)),
  Std_Dev = round(data_num_pca$sdev, 2),
  Prop_Variance = round(exp_var, 2),
  Cum_Variance = round(data_ex_vratio, 2)
)

pca_tab <- tableGrob(pca_summary)
grid.newpage()
grid.draw(pca_tab)

PC1 <- data_num_pca$x[,1]
cor.test(PC1, data_c$LN_IC50) #show PC 1 confidence interval roughly between 0.746 and 0.750

plot(
  PC1,
  data_c$LN_IC50,
  pch = 19,
  xlab = "PC1",
  ylab = "LN_IC50",
  main = "PC1 vs LN_IC50"
)
abline(lm(data_c$LN_IC50 ~ PC1), col = "red")
```

```{r}
# the final cleaned, reduced, and transformed data set is data_f.
data_f <- data_c %>% select(-AUC)
str(data_f)
```

#SAM

#Since the models took an extremely long time to train due to the caret package
#naturally encoding all your categorical variables to be dummy variables, it was computationally expensive.
#Instead predictors like CELL_LINE_NAME and DRUG_NAME are removed from the model to decrease computation time
#and they did not provide predictive power

#Seed is set. Split the data on cell line

```{r}
seed <- 756
set.seed(seed)

#Target variable
IC50 <- data_f$LN_IC50

#Having just predictors in the data set
predictors <- data_f %>% select(-LN_IC50)

#Get unique cell lines
cell_lines_names <- unique(predictors$CELL_LINE_NAME)

#20 percent of the cell line names as left as test (splitting the data based on cell line name)
test_set_samples <- sample(cell_lines_names, size = floor(0.2 * length(cell_lines_names)))

testingRows <- predictors$CELL_LINE_NAME %in% test_set_samples

testPredictors <- predictors[testingRows,]
trainPredictors <- predictors[!testingRows,]

trainIC50 <- IC50[!testingRows]
testIC50  <- IC50[testingRows]

dim(trainPredictors)
dim(testPredictors)

length(trainIC50)
length(testIC50)
```

#Encode -- One hot encode all the predictor values manually (not through caret model)
#that will be used for our model predictions

```{r}
#Make variables from character to factor type for encoding 
trainPredictors$TARGET <- factor(trainPredictors$TARGET)
testPredictors$TARGET <- factor(testPredictors$TARGET,
                                levels = levels(trainPredictors$TARGET))

trainPredictors$TARGET_PATHWAY <- factor(trainPredictors$TARGET_PATHWAY)
testPredictors$TARGET_PATHWAY  <- factor(testPredictors$TARGET_PATHWAY,
                                         levels = levels(trainPredictors$TARGET_PATHWAY))

trainPredictors$Microsatellite.instability.Status..MSI. <- 
    factor(trainPredictors$Microsatellite.instability.Status..MSI.)

testPredictors$Microsatellite.instability.Status..MSI.  <- 
    factor(testPredictors$Microsatellite.instability.Status..MSI.,
           levels = levels(trainPredictors$Microsatellite.instability.Status..MSI.))
```

#Encoding TARGET and TARGET_PATHWAY variables

```{r}
target_encode_tr  <- model.matrix(~ TARGET - 1, data = trainPredictors)
target_encode_tst <- model.matrix(~ TARGET - 1, data = testPredictors)

pathway_encode_tr  <- model.matrix(~ TARGET_PATHWAY - 1, data = trainPredictors)
pathway_encode_tst <- model.matrix(~ TARGET_PATHWAY - 1, data = testPredictors)

#Line up columns for model testing
target_encode_tst  <- target_encode_tst[,  colnames(target_encode_tr),  drop = FALSE]
pathway_encode_tst <- pathway_encode_tst[, colnames(pathway_encode_tr), drop = FALSE]
```

#Encoding TCGA_DESC variable

```{r}
tissue_dummies <- dummyVars(~ TCGA_DESC, data = trainPredictors)
tissue_encode_tr  <- predict(tissue_dummies, trainPredictors)
tissue_encode_tst <- predict(tissue_dummies, testPredictors)
```

#Same encoding process for GDSC Tissue Descriptor 2 as TCGA_DESC

```{r}
gdsctd2_dummies <- dummyVars(~ GDSC.Tissue.descriptor.2, data = trainPredictors)
gdsctd2_encode_tr  <- predict(gdsctd2_dummies, trainPredictors)
gdsctd2_encode_tst <- predict(gdsctd2_dummies, testPredictors)
```

#Encoding continuous variables

```{r}
MSI_encode_tr <- as.numeric(trainPredictors$Microsatellite.instability.Status..MSI.)
MSI_encode_tst  <- as.numeric(testPredictors$Microsatellite.instability.Status..MSI.)

Zscore_encode_tr <- trainPredictors$Z_SCORE
Zscore_encode_tst <- testPredictors$Z_SCORE
```

```{r}
trainEncodedPredictors <- data.frame(
  target_encode_tr,
  pathway_encode_tr,
  tissue_encode_tr,
  gdsctd2_encode_tr,
  MSI_encode_tr,
  Zscore_encode_tr
)

testEncodedPredictors <- data.frame(
  target_encode_tst,
  pathway_encode_tst,
  tissue_encode_tst,
  gdsctd2_encode_tst,
  MSI_encode_tst,
  Zscore_encode_tst
)
```

#Have train and test sets have the same number of predictors

```{r}
#Align columns for train and test
common_cols <- union(colnames(trainEncodedPredictors),
                     colnames(testEncodedPredictors))

#Add missing columns to training set
missing_train <- setdiff(common_cols, colnames(trainEncodedPredictors))
trainEncodedPredictors[, missing_train] <- 0

#Add missing columns to test set
missing_test <- setdiff(common_cols, colnames(testEncodedPredictors))
testEncodedPredictors[, missing_test] <- 0

#Reorder columns 
trainEncodedPredictors <- trainEncodedPredictors[, common_cols]
testEncodedPredictors  <- testEncodedPredictors[, common_cols]
```

```{r}
#Set up cross validation 
threefoldCV <- trainControl(method = "cv", number = 3, allowParallel = TRUE)
#repeatedSplits <- trainControl(method = "repeatedcv", number = 5, repeats = 3, allowParallel = TRUE)
```

```{r}
#Getting rid of Z-score
trainEncodedPredictors <- trainEncodedPredictors %>% select(-Zscore_encode_tr)
testEncodedPredictors  <- testEncodedPredictors  %>% select(-Zscore_encode_tr)

#Sanity Check
dim(trainEncodedPredictors)
dim(testEncodedPredictors)
length(trainIC50)
```

#Linear Regression use as a baseline

```{r}
#Before model training we need to use parallel processing so the CPU can run models more quickly
num_cores <- detectCores() - 1             # Leave 1 core for background tasks
cl <- makeCluster(num_cores)
registerDoParallel(cl)


set.seed(seed)
lin_reg_baseline <- train(trainEncodedPredictors, trainIC50,
                 method = "lm",
                 trControl = threefoldCV)
lin_reg_baseline

#Stop the parallel cluster when models are finished training
stopCluster(cl)
registerDoSEQ()
```

```{r}
lin_reg_Imp <- varImp(lin_reg_baseline, scale = FALSE)
plot(lin_reg_Imp, top = 25)
```

#Important predictors that will be used for our other training models

```{r}
important_feature_values <- lin_reg_Imp$importance$Overall
length(important_feature_values)
names(important_feature_values) <- rownames(lin_reg_Imp$importance)
top25 <- sort(important_feature_values, decreasing = TRUE)[1:25]
top25
top25_cols <- names(sort(important_feature_values, decreasing = TRUE))[1:25]
top25_cols
```


#Train all features with reduced predictors

```{r}
trainReduced <- trainEncodedPredictors[, top25_cols]
testReduced  <- testEncodedPredictors[, top25_cols]
```

#Reduced Rows for computation speed

```{r}
set.seed(seed)
row_reduction <- 0.15

trainEncoded_df <- trainEncodedPredictors %>%
  mutate(
    CELL_LINE_NAME = trainPredictors$CELL_LINE_NAME,
    IC50 = trainIC50
  )
trainReducedRows_df <- trainEncoded_df %>%
  group_by(CELL_LINE_NAME) %>%
  sample_frac(row_reduction) %>%
  ungroup()

trainReducedRows <- trainReducedRows_df[, top25_cols]
trainIC50ReducedRows <- trainReducedRows_df$IC50
```

```{r}
testReduced2 <- testReduced
testIC50Reduced <- testIC50
```


```{r}
#Before model training we need to use parallel processing so the CPU can run models more quickly
num_cores <- detectCores() - 1             # Leave 1 core for background tasks
cl <- makeCluster(num_cores)
registerDoParallel(cl)

lin_reg <- train(trainReducedRows, trainIC50ReducedRows,
                 method = "lm",
                 trControl = threefoldCV)
lin_reg

#Stop the parallel cluster when models are finished training
stopCluster(cl)
registerDoSEQ()
```


#KNN regression

```{r}
set.seed(seed)
#Before model training we need to use parallel processing so the CPU can run models more quickly
num_cores <- detectCores() - 1             # Leave 1 core for background tasks
cl <- makeCluster(num_cores)
registerDoParallel(cl)

knn <- train(trainReducedRows, trainIC50ReducedRows,
             method = "kknn",
             tuneGrid = data.frame(
             kmax = seq(1, 11, by = 2), 
             distance = 2,            
             kernel = "rectangular"           
             ),
             preProcess = c("center", "scale"),
             trControl = threefoldCV)
knn

#Stop the parallel cluster when models are finished training
stopCluster(cl)
registerDoSEQ()
```
```{r}
plot(knn)
```


#SVM

```{r}
set.seed(seed)
#Before model training we need to use parallel processing so the CPU can run models more quickly
num_cores <- detectCores() - 1             # Leave 1 core for background tasks
cl <- makeCluster(num_cores)
registerDoParallel(cl)

svmGrid <- expand.grid(sigma = 10^seq(-2, 1, length = 4),
                       C = 10^seq(-2,1, length = 4))
SVM <- train(trainReducedRows, trainIC50ReducedRows,
             method = "svmRadial",
             tuneGrid = svmGrid,
             preProcess = c("center", "scale"),
             trControl = threefoldCV) 
SVM
#Stop the parallel cluster when models are finished training
stopCluster(cl)
registerDoSEQ()

```

```{r}
target_microtibule_stabiliser <- tibble(target_microtibule_stabiliser = seq(min(trainReducedRows$TARGETMicrotubule.stabiliser),
                                     max(trainReducedRows$TARGETMicrotubule.stabiliser),
                                     length.out = 200))
```

```{r}
make_prediction_grid <- function(var_name, train_reduced_data){
  
  # Since model has 25 predictors get the medians of the other 24
  #predictors to hold them constant so we can visualize the change of
  #target.microtibule_stabiliser 
  median_predictors <- as.list(apply(train_reduced_data, 2, median))
  
  # Dummy variable can only be 0 or 1
  dummy_vals <- c(0, 1)
  
  # Make a tibble data.frame with 200 rows and the median predictors
  SVM_tibble <- tibble()
  for (val in dummy_vals) {
    row <- median_predictors
    row[[var_name]] <- val
    SVM_tibble <- bind_rows(SVM_tibble, as_tibble(row))
  }
  
  SVM_tibble
}

prediction_grid <- make_prediction_grid(
  "TARGETMicrotubule.stabiliser",
  trainReducedRows
)
```

```{r}
SVM_explore <- function(cost, sigma, grid, train_data, train_y) {
  
  # Fit an SVM with the given cost and sigma
  svmFit <- ksvm(
    x = as.matrix(train_data),
    y = train_y,
    kernel = "rbfdot",
    C = cost,
    kpar = list(sigma = sigma),
    scaled = TRUE
  )
  
  # Predict on the grid
  preds <- predict(svmFit, as.matrix(grid))
  
  tibble(
    TARGETMicrotubule.stabiliser = grid$TARGETMicrotubule.stabiliser,
    IC50_pred = preds,
    cost = cost,
    sigma = sigma
  )
}

```

```{r}
best_sigma <- SVM$bestTune$sigma
cost_values <- 10^seq(-2, 1, length = 4)

cost_curves <- map_dfr(cost_values, function(c) {
  SVM_explore(c, best_sigma, prediction_grid, trainReducedRows, trainIC50ReducedRows)
})
```

```{r}
ggplot(cost_curves,
       aes(x = TARGETMicrotubule.stabiliser, y = IC50_pred, color = factor(cost))) +
  geom_line(size = 1) +
  labs(color = "Cost")
```

```{r}
best_cost <- SVM$bestTune$C
sigma_values <- 10^seq(-2, 1, length = 4)

sigma_curves <- map_dfr(sigma_values, function(s) {
  SVM_explore(best_cost, s, prediction_grid, trainReducedRows, trainIC50ReducedRows)
})
```

```{r}
ggplot(sigma_curves,
       aes(x = TARGETMicrotubule.stabiliser, y = IC50_pred, color = factor(sigma))) +
  geom_line(size = 1) +
  labs(color = "Sigma")
```

```{r}
trainReducedRows_num <- data.frame(lapply(trainReducedRows, function(x) as.numeric(as.character(x))))
testReduced2_num <- data.frame(lapply(testReduced2, function(x) as.numeric(as.character(x))))
```

#MARS

```{r}
set.seed(seed)
#Before model training we need to use parallel processing so the CPU can run models more quickly
num_cores <- detectCores() - 1             # Leave 1 core for background tasks
cl <- makeCluster(num_cores)
registerDoParallel(cl)

MARSGrid <- expand.grid(degree = 1:2, nprune = 10:35)
MARS <- train(trainReducedRows_num, trainIC50ReducedRows,
              method ="earth",
              tuneGrid = MARSGrid,
              trControl = threefoldCV)
MARS

#Stop the parallel cluster when models are finished training
stopCluster(cl)
registerDoSEQ()
```

```{r}
plot(MARS)
```


#Random Forest

```{r}
set.seed(seed)
#Before model training we need to use parallel processing so the CPU can run models more quickly
num_cores <- detectCores() - 1             # Leave 1 core for background tasks
cl <- makeCluster(num_cores)
registerDoParallel(cl)

mtryGrid <- data.frame(
            mtry = c(1:10, 12, 15, 20))
rf <- train(trainReducedRows, trainIC50ReducedRows,
                method = "rf",
                tuneGrid = mtryGrid,
                ntree = 500,
                importance = TRUE,
                trControl = threefoldCV
            )
rf

#Stop the parallel cluster when models are finished training
stopCluster(cl)
registerDoSEQ()
```

```{r}
plot(rf)
```

```{r}
#Keep only the test rows whose cell lines were used in the reduced training set.
testResults_lin_regr <- predict(lin_reg, testReduced2)
testResults_knn_regr <- predict(knn, testReduced2)
testResults_SVM <- predict(SVM, testReduced2)
testResults_MARS <- predict(MARS, testReduced2_num)
testResults_rf <- predict(rf, testReduced2)

#Use postSample to evaluate the RMSE, Rsquared, and MAE
lin_regr_metrics  <- postResample(testResults_lin_regr, testIC50Reduced)
knn_regr_metrics  <- postResample(testResults_knn_regr, testIC50Reduced)
SVM_metrics  <- postResample(testResults_SVM, testIC50Reduced)
MARS_metrics <- postResample(testResults_MARS, testIC50Reduced)
rf_metrics   <- postResample(testResults_rf, testIC50Reduced)
```


```{r}
#model results on y_test 
model_results <- cbind(
  Linear_Regression = lin_regr_metrics,
  Knn_Regression = knn_regr_metrics,
  SVM = SVM_metrics,
  MARS = MARS_metrics,
  Random_Forest = rf_metrics
)
model_results
```

