---
title: "Genomics of Drug Sensitivity in Cancer (GDSC) Project"
author: "Sam Scott, Megan Lee, Titus Sun"
date: "2026-06-01"
output: html_document
---

```{r warning=FALSE, message=FALSE}
library(caret)
library(tidyverse)
library(kernlab)
library(dplyr)
library(tibble)
library(mlbench)
library(randomForest)
```

# a way to read the dataset off of github so no need for local machine calls
```{r}
data_url <- "https://raw.githubusercontent.com/sascott9655/GDSC-Project/refs/heads/main/GDSC_DATASET.csv"
data  <- read.csv(data_url)
```

```{r}
str(data)
head(data)
dim(data)
```

```{r}
# cleaning the data

# verifying the COSMIC_ID numbers are unique to each CELL_LINE_NAME

data %>%
  group_by(COSMIC_ID) %>%
  summarise(n_unique = n_distinct(CELL_LINE_NAME)) %>%
  filter(n_unique > 1)

# verifying the DRUG_ID numbers are unique to each DRUG_NAME

data %>%
  group_by(DRUG_ID) %>%
  summarise(n_unique = n_distinct(DRUG_NAME)) %>%
  filter(n_unique > 1)

# assessing the redundancy of the TCGA columns

tcga_match <- data$TCGA_DESC == data$Cancer.Type..matching.TCGA.label.
table(tcga_match, useNA = "ifany")

#tcga_miss <- data[data$TCGA_DESC != data$Cancer.Type..matching.TCGA.label., ]
#dim(tcga_miss)
#sum(tcga_miss$Cancer.Type..matching.TCGA.label. == "", na.rm = TRUE)
#sum(
  #!tcga_match &
  #!is.na(data$TCGA_DESC) &
  #!is.na(data$Cancer.Type..matching.TCGA.label.) &
  #data$TCGA_DESC != "" &
  #data$Cancer.Type..matching.TCGA.label. != ""
#)

sum(data$Cancer.Type..matching.TCGA.label. == "")
sum(data$TCGA_DESC == "")
sum(data$Cancer.Type..matching.TCGA.label. == "" &
      data$TCGA_DESC == "")

#tcga_miss2 <- data[
    #!tcga_match &
    #!is.na(data$TCGA_DESC) &
    #!is.na(data$Cancer.Type..matching.TCGA.label.) &
    #data$TCGA_DESC != "" &
    #data$Cancer.Type..matching.TCGA.label. != "",
  #]
#head(tcga_miss2, 20)

# removing irrelevant and redundant columns

data2 <- data %>%
  select(-COSMIC_ID, -DRUG_ID, -Screen.Medium, -Growth.Properties, -CNA, -Gene.Expression, -Methylation)
str(data2)

# looking for na and blank values

data2 %>%
  summarise(across(everything(), ~sum(is.na(.))))%>%
  t()

blank <- sapply(data2, function(x) sum(x == "", na.rm = TRUE))
as_tibble(list(
  column = names(blank),
  empty_count = as.vector(blank)
))

# reference Cancer.Type..matching.TCGA.label. to enter values for TCGA_DESC

unique(data2$Cancer.Type..matching.TCGA.label.)
data2 <- data2 %>%
  mutate(
    TCGA_DESC = case_when(
      TCGA_DESC == "" & Cancer.Type..matching.TCGA.label. == "" ~ "UNCLASSIFIED1",
      TCGA_DESC == "" ~ Cancer.Type..matching.TCGA.label.,
      TRUE ~ TCGA_DESC
    )
  )

blank <- sapply(data2, function(x) sum(x == "", na.rm = TRUE))
as_tibble(list(
  column = names(blank),
  empty_count = as.vector(blank)
))

# reference TCGA_DESC to enter GDSC.Tissue.descriptor.1 values for blank values in GDSC.Tissue.descriptor.2

length(unique(data2$GDSC.Tissue.descriptor.1))
length(unique(data2$GDSC.Tissue.descriptor.2))
length(unique(data2$TCGA_DESC))

gdsc_miss <- data2 %>%
  filter(GDSC.Tissue.descriptor.1 != "", !is.na(GDSC.Tissue.descriptor.1)) %>%
  group_by(TCGA_DESC, GDSC.Tissue.descriptor.1) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(TCGA_DESC) %>%
  slice_max(n, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(TCGA_DESC, GDSC.Tissue.descriptor.1)

gdsc_miss <- gdsc_miss %>%
  rename(gdsc_map = GDSC.Tissue.descriptor.1)

data2 <- data2 %>%
  left_join(gdsc_miss, by = "TCGA_DESC") %>%
  mutate(
    GDSC.Tissue.descriptor.2 = case_when(
      GDSC.Tissue.descriptor.2 != "" ~ GDSC.Tissue.descriptor.2,
      GDSC.Tissue.descriptor.1 != "" ~ GDSC.Tissue.descriptor.1,
      !is.na(gdsc_map) ~ gdsc_map,
      TRUE ~ "UNCLASSIFIED1"
    )
  )

blank <- sapply(data2, function(x) sum(x == "", na.rm = TRUE))
as_tibble(list(
  column = names(blank),
  empty_count = as.vector(blank)
))

# imputing blank MSI values

unique(data2$Microsatellite.instability.Status..MSI.)

tcga_msi <- data2 %>%
  group_by(TCGA_DESC) %>%
  summarise(
    n_groups = n_distinct(Microsatellite.instability.Status..MSI.),
    groups = paste(unique(Microsatellite.instability.Status..MSI.), collapse = ", "),
    .groups = "drop"
  )
tcga_msi

table(data2$Microsatellite.instability.Status..MSI.)

msi_miss <- data2 %>%
  filter(Microsatellite.instability.Status..MSI. != "", !is.na(Microsatellite.instability.Status..MSI.)) %>%
  group_by(TCGA_DESC, Microsatellite.instability.Status..MSI.) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(TCGA_DESC) %>%
  slice_max(n, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(TCGA_DESC, Microsatellite.instability.Status..MSI.)

msi_miss <- msi_miss %>%
  rename(msi_map = Microsatellite.instability.Status..MSI.)

data2 <- data2 %>%
  left_join(msi_miss, by = "TCGA_DESC") %>%
  mutate(
    Microsatellite.instability.Status..MSI. = coalesce(
      na_if(Microsatellite.instability.Status..MSI., ""),
      msi_map,
      "UNCLASSIFIED1"
    )
  )

blank <- sapply(data2, function(x) sum(x == "", na.rm = TRUE))
as_tibble(list(
  column = names(blank),
  empty_count = as.vector(blank)
))

# imputing blank TARGET values

sum(data2$TARGET == "" & data2$TARGET_PATHWAY != "Unclassified")
target_miss <- data2 %>%
  filter(TARGET == "", TARGET_PATHWAY != "Unclassified")
unique(target_miss$TARGET_PATHWAY)



















